---
title: "Final Report"
author: "Group 1"
date: "2024-09-18"
output: 
    html_document:
        self_contained: TRUE
        output_file: "Data_Deployment.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
### Setup ###
# Loading Libraries
library(dplyr)
library(tidyverse)
library(readr)
library(here)
library(tidyr)
library(broom)
library(kableExtra)
```

# Research Motivation

User ratings are found to be a key success factor of series and help production companies gain financial success (Hsu et al., 2014). As such, it is important to determine what factors influence these ratings, in order to develop series that are most likely to receive high ratings. The length of an individual episode is found to have a significant effect on the rating on that episode (Danaher et al., 2011). However, how the length of the complete series influences the rating of the episodes remains unexplored. This study fills the gap in the literature by exploring the following research question:

What is the effect of the length of a TV show on the average rating of the TV show episodes?

# Research Method

The chosen research method for the mentioned research question is a regression analysis. This research method will be suitable to investigate the relation between the following variables: the length of a TV show and the episode ratings. Performing a regression analysis allows us to determine if the length of a TV show significantly affects the ratings of its episodes and to learn more about the direction of the relationship.

To perform the regression analysis we will make use of the following data sets: 1. title.episode.tsv.gz a. To find the series and the episodes 2. title.ratings.tsv.gz a. Average rating of individual users combined 3. title.basics.tsv.gz a. Genre for cross reference

# Results

The aim of this study was to explore the relationship between TV show length and episode rating. A regression analysis, supported by a plot visualization (both can be seen below), examined how runtime impacts viewer evaluations.

The results show a negative relationship between series runtime and ratings. The intercept is 7.864, indicating that a show with 0 runtime would have an estimated rating of 7.86. The coefficient for runtime is -0.0000142, meaning each additional minute slightly decreases the rating by 0.0000142 points.

The result is statistically significant, with a p-value of 0, confirming that the relationship is unlikely to be due to chance. While the effect is small, longer runtimes do to result in lower ratings. Both the regression analysis and plot visualization can be seen below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Results:
### Input ###
#Creating the data df
data <- read_csv(here("gen", "data_preparation", "output", "cleaned_data.csv"))

### Transformation/Output ###

### Linear regression model
# Create function to do the regression analysis and make a table for the report
run_regression_analysis <- function(data) {
  model <- lm(rating ~ series_runtime, data = data)
  
  # Extract a clean data frame of the results
  tidy_model <- tidy(model)
  
  # Generate a proper table for in the report
  tidy_table <- kable(
    tidy_model, 
    format = "html",
    caption = "Regression Analysis: Rating vs Series Runtime",
    col.names = c("Term", "Estimate", "Standard Error", "Statistic", "P-value")
  ) %>%
    kable_styling(full_width = FALSE, position = "center")
  
    return(tidy_table)
}
run_regression_analysis(data)

#### Relationship between TV show runtime and ratings
ggplot(data, aes(x = series_runtime, y = rating)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(title = "Relationship Between TV show Runtime and Ratings", x = "TV show Runtime (minutes)", y = "Rating")

```

# Theoretical / Practical Implications

## Theoretical implications

This study contributes to academic understanding by addressing how structural elements, such as TV show runtime, influence viewer ratings. The finding that longer TV shows tend to receive lower ratings suggests that increased runtime may allow for diminishing quality over time. Furthermore, viewers might lose attention and become less engaged with the series, thereby lowering viewer satisfaction. This finding extends media theory by offering new insights into how content structure — specifically series length — impacts audience engagement. Additionally, the observed decline in audience satisfaction across seasons highlights the difficulty of maintaining viewer interest over time, contributing to broader discussions about media consumption and how series length can influence ratings and viewer satisfaction.

## Practical implications

This study provides insights for the TV and media production industry. The finding that longer TV shows lead to lower ratings can help production companies optimize the structure of their series by adjusting series length to meet viewer preferences and improve engagement.They are recommended to not stretch out TV series, but focus on maintaining the quality of the series, thereby hoping to limit this negative effect on ratings. Furthermore, streaming services that release entire seasons at once could use these findings to better predict viewer behavior and ratings based on series length, helping them tailor their release strategies to maximize ratings and engagement. Since ratings are linked to financial success, aligning series length with audience expectations could help production companies boost profitability by increasing viewer satisfaction and long-term engagement.

# Conclusion

The aim of this study was to explore the relationship between the length of a TV show and the average rating of the TV show episodes. Using a linear regression, supported by a scatter plot visualization, the analysis examined how TV show runtime impacted viewer evaluations.

The results of the regression analysis reveal a significant negative effect of TV show length on episode rating. The plot visualizes this effect. As such, the longer a TV show runs, the lower the ratings of the episodes will be. A shorter TV show runtime thus leads to a higher rating.
